---
title: "Chi Square"
author: "Kübra MAHMUT ÖNALAN"
output:
  word_document: default
  html_notebook: default
---

# Chi-Square  

There are two types of Chi-Square tests that are commonly used:

Chi-Square Goodness of Fit Test: Used to determine whether or not a categorical variable follows a hypothesized distribution.
Chi-Square Test of Independence: Used to determine whether or not there is a significant association between two categorical variables.


# Download packages 

```{r}
library(readxl)  
library (dplyr)
library(vcd)
library(grid)
```

# Set directory 

```{r}

setwd("")
```

#Import and Organize Dataset

```{r}
df_all <- read_excel("")

```

```{r}

df_chi <- df_all %>%
  
  select(CINSIYET,PSİ_YARDIM_GECMİS, PSİ_YARDIM_SİMDİ)


colnames(df_chi)[1] <- "Gender"
colnames(df_chi)[2] <- "Previous"
colnames(df_chi)[3] <- "Now"



df_chi$Gender[df_chi$Gender == 1] <- 2 # male
df_chi$Gender[df_chi$Gender == 0] <- 1 # female



df_chi$Previous[df_chi$Previous == 1] <- 2 # No
df_chi$Previous[df_chi$Previous == 0] <- 1 # Yes


df_chi$Now[df_chi$Now == 1] <- 2 # No
df_chi$Now[df_chi$Now == 0] <- 1 # Yes


df_chi$Gender <- factor(df_chi$Gender ,levels = c(1,2), labels = c("Female", "Male"))
df_chi$Previous <- factor(df_chi$Previous ,levels = c(1,2), labels = c("Yes", "No"))
df_chi$Now<- factor(df_chi$Now ,levels = c(1,2), labels = c("Yes", "No"))



df_chi



```


# Table Gender and Previous

```{r}

cqi_table1  <-  df_chi  %>% select(Previous, Gender)

cqi_table1 <-  table(cqi_table1)

cqi_table1


```


# Table Gender and Now

```{r}

cqi_table2  <-  df_chi  %>% select(Now, Gender)

cqi_table2<-  table(cqi_table2)

cqi_table2



```

#Chi- Square Analysis - Chi-Square Test of Independence 


```{r}


chisq1 <- chisq.test(cqi_table1)

chisq1

chisq1$expected

chisq1$observed

```


```{r}

chisq2 <- chisq.test(cqi_table2)

chisq2

chisq2$expected

chisq2$observed

```


"Psychiatric/ Psychological Treatment Previous"
 "Psychiatric/ Psychological Treatment Now"
 

```{r}

mosaic( Gender ~  Previous, data = cqi_table1 ,main = "Psychiatric/ Psychological Treatment Previous", direction=c("v","h"), shade = TRUE, gp = gpar(fill=c('lightblue', 'pink')))

```
 
 

```{r}



mosaic( Gender ~  Now, data = cqi_table2,main = "Psychiatric/ Psychological Treatment Now", direction=c("v","h"), shade = TRUE, gp = gpar(fill=c('lightblue', 'pink')))
```


# Report 

we use the following general structure to report the results of a Chi-Square Test of Independence in APA format:

A Chi-Square Test of Independence was performed to assess the relationship between [variable 1] and [variable 2].

There [was or was not] a significant relationship between the two variables, X2(df, N) = [X2 value], p = [p-value].

