---
title: " Mediation Analysis"
output:
  author 
---

# Checking Normal Distrubution + Visulations 
# Corelation Analysis  + Visulations 
# Multiple Regression - Regression Model Diagnotistic 
# Mediator Analysis 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) 
library (dplyr) 
library(gtsummary)
library(flextable)
library (rstatix) 
library(jmv) 
library(ggplot2)
library(ggpubr) 
library(apaTables) 
library (GGally)
```

```{r}

setwd("")

```

```{r}
df_all <- read_excel("")

colnames(df_all)

```


## 1. Descriptive Analysis and Checking Normal Distirubiton  

# 1.1. Descriptive Analysis 
```{r}

df_descriptive <- df_all %>%
 dplyr::select(YAS, CINSIYET,M_DURUM, E_DURUM,PSİ_YARDIM_GECMİS, PSİ_YARDIM_SİMDİ)


colnames(df_descriptive)[1] <- "Age"
colnames(df_descriptive)[2] <- "Gender"
colnames(df_descriptive)[3] <- "Relationship_Status"
colnames(df_descriptive)[4] <- "Educational_Level"
colnames(df_descriptive)[5] <- "PPT_Previous"
colnames(df_descriptive)[6] <- "PPT_Now"



df_descriptive$Gender[df_descriptive$Gender == 1] <- 2 # male
df_descriptive$Gender[df_descriptive$Gender == 0] <- 1 # female


df_descriptive$Relationship_Status[df_descriptive$Relationship_Status == 2] <- 3 #Divorced
df_descriptive$Relationship_Status[df_descriptive$Relationship_Status == 1] <- 2 # Married
df_descriptive$Relationship_Status[df_descriptive$Relationship_Status == 0] <- 1 # Single


df_descriptive$PPT_Previous[df_descriptive$PPT_Previous == 1] <- 2 # No
df_descriptive$PPT_Previous[df_descriptive$PPT_Previous == 0] <- 1 # Yes


df_descriptive$PPT_Now[df_descriptive$PPT_Now == 1] <- 2 # No
df_descriptive$PPT_Now[df_descriptive$PPT_Now == 0] <- 1 # Yes




df_descriptive$Gender <- factor(df_descriptive$Gender ,levels = c(1,2), labels = c("Female", "Male"))
df_descriptive$Relationship_Status <- factor(df_descriptive$Relationship_Status ,levels = c(1,2,3), labels = c("Single", "Married", "Divorced"))
df_descriptive$Educational_Level <- factor(df_descriptive$Educational_Level ,levels = c(1,2), labels = c("High School", "University-Graduate"))
df_descriptive$PPT_Previous <- factor(df_descriptive$PPT_Previous ,levels = c(1,2), labels = c("Yes", "No"))
df_descriptive$PPT_Now<- factor(df_descriptive$PPT_Now ,levels = c(1,2), labels = c("Yes", "No"))

df_descriptive

```

```{r}

tabledemog<- tbl_summary(df_descriptive,  by = "Gender",
                     statistic = list(all_continuous() ~ "{mean} ({sd}) {min} - {max}",
                                      all_categorical() ~ "{n}({p}%)"),
                     type = list(PPT_Previous ~ "categorical", PPT_Now ~ "categorical"),
                     label =list(Relationship_Status~ "Relationship Status",Educational_Level~ "Educational Level",
                                 PPT_Previous ~ "Psychiatric/Psychological Treatment (Previous)",
                                 PPT_Now ~ "Psychiatric/Psychological Treatment (Now)"),
                     missing_text = "Missing")%>%
  modify_header(label = "**Variables**",all_stat_cols() ~ "**{level}**<br>N = {n} (%{style_percent(p)})") %>%
  add_overall(col_label = "**Sample**") %>%
  add_n() %>%
  bold_labels()

tabledemog

```

```{r}

df_mainvariables <- df_all %>%
  dplyr::select(CINSIYET,BEEtop, RTOTOP, BIFOTOP, BDETOP)

df_mainvariables


colnames(df_mainvariables)[1] <- "Gender"
colnames(df_mainvariables)[2] <- "CogFlex"
colnames(df_mainvariables)[3] <- "Rumination"
colnames(df_mainvariables)[4] <- "Mindfulness"
colnames(df_mainvariables)[5] <- "Depression"

df_mainvariables

df_mainvariables$Gender[df_mainvariables$Gender == 1] <- 2 # male
df_mainvariables$Gender[df_mainvariables$Gender == 0] <- 1 # female

df_mainvariables$Gender <- factor(df_mainvariables$Gender ,levels = c(1,2), labels = c("Female", "Male"))

df_mainvariables


```

```{r}

sum <-df_mainvariables %>% descriptives(hist = FALSE,
             violin = FALSE, 
             dot = FALSE, dotType = "jitter", 
             qq = FALSE,
             n = TRUE, 
             missing = TRUE, 
             mean = TRUE, 
             median = TRUE,
             mode = TRUE, 
             sum = TRUE, 
             sd = TRUE, 
             variance = TRUE,
             range = FALSE, 
             min = TRUE, 
             max = TRUE, 
             se = FALSE,
             skew = TRUE, kurt = TRUE)

sum

```


```{r}

tablemain <- tbl_summary(df_mainvariables,
            by= "Gender",
          type = all_continuous() ~ "continuous2",
          statistic = all_continuous() ~ c("{N_nonmiss}",
                           "{mean}", 
                           "{sd}",
                           "{min}",
                           "{max}"),
          label =list(CogFlex~ "Cognitive Flexibility"),
          digits = list(everything() ~ c(2))) %>%
  modify_header(all_stat_cols() ~ "**{level}**<br>N = {n} (%{style_percent(p)})") %>%
  add_overall(col_label = "**Sample**") 

tablemain

```

# 1.1. Normal Distirubiton (Histogram + Density and QQ Plots) 

# 1.1.1. Cognitive Flexibility 

```{r}

P1 <- ggplot(data = df_mainvariables, aes(CogFlex)) +   
  geom_histogram(aes(y = ..density..),bins = 15, position = "identity", color ="#D55E00", fill = "#D55E00",  linetype="dashed",  alpha = 0.5) + 
  ggtitle("Cognitive Flexibility Histogram and Density Graph") +
  geom_vline(aes(xintercept=mean(CogFlex)),
             color="#000000", linetype="dashed", size=0.75) +
  xlab("Cognitive Flexibility")+
  ylab ("Density") + 
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5)) +
  stat_function(fun = dnorm,
                args = list(mean = mean(df_mainvariables$CogFlex),
                            sd = sd(df_mainvariables$CogFlex)),
                col = "#000000",
                size = 1) + 
  theme_classic()

P1

```

```{r}

P2 <- ggplot(df_mainvariables, aes(sample=CogFlex))+
  stat_qq(color= "#D55E00", shapetype = "point", alpha = 0.5, bins = 15)+
  ggtitle("Cognitive Flexibility QQ PLot") +
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5))+
  theme_classic()

P2



```

#1.1.2. Rumination 

```{r}
P3 <- ggplot(data = df_mainvariables, aes(Rumination)) +   
  geom_histogram(aes(y = ..density..),bins = 15, position = "identity", color ="#D55E00", fill = "#D55E00",  linetype="dashed",  alpha = 0.5) + 
  ggtitle("Rumination Histogram and Density Graph") +
  geom_vline(aes(xintercept=mean(Rumination)),
             color="#000000", linetype="dashed", size=0.75) +
  xlab("Rumination")+
  ylab ("Density") + 
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5)) +
  stat_function(fun = dnorm,
                args = list(mean = mean(df_mainvariables$Rumination),
                            sd = sd(df_mainvariables$Rumination)),
                col = "#000000",
                size = 1) + 
  theme_classic()

P3

```

```{r}

P4 <- ggplot(df_mainvariables, aes(sample=Rumination))+
  stat_qq(color= "#D55E00", shapetype = "point", alpha = 0.5, bins = 15)+
  ggtitle("Rumination QQ PLot") +
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5))+
  theme_classic()
P4


```


#1.1.3. Mindfulness


```{r}


P5 <- ggplot(data = df_mainvariables, aes(Mindfulness)) +  
  geom_histogram(aes(y = ..density..),bins = 15,
                 color ="#D55E00", fill = "#D55E00",
                 linetype="dashed",  alpha = 0.5) + 
  ggtitle("Mindfulness Histogram and Density Graph") +
  geom_vline(aes(xintercept=mean(Mindfulness)),
             color="#000000", linetype="dashed", size=1) +
  xlab("Mindfulness")+
  ylab ("Density") + 
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5)) +
  stat_function(fun = dnorm,
                args = list(mean = mean(df_mainvariables$Mindfulness),
                            sd = sd(df_mainvariables$Mindfulness)),
                col = "#000000",
                size = .80) + 
   
  theme_classic()

P5

```



```{r}

P6 <- ggplot(df_mainvariables, aes(sample=Mindfulness))+
  stat_qq(color= "#D55E00", shapetype = "point", alpha = 0.5, bins = 15)+
  ggtitle("Mindfulness QQ Plot") +
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5))+
  theme_classic()
P6

```

#1.1.4. Depression 

```{r}

P7 <- ggplot(data = df_mainvariables, aes(Depression)) +   
  geom_histogram(aes(y = ..density..),bins = 15, position = "identity", color ="#D55E00", fill = "#D55E00",  linetype="dashed",  alpha = 0.5) + 
  ggtitle("Depression Histogram and Density Graph") +
  geom_vline(aes(xintercept=mean(Depression)),
             color="#000000", linetype="dashed", size=0.75) +
  xlab("Depression")+
  ylab (".") + 
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5)) +
  stat_function(fun = dnorm,
                args = list(mean = mean(df_mainvariables$Depression),
                            sd = sd(df_mainvariables$Depression)),
                col = "#000000",
                size = 1) + 
  theme_classic()

P7

```


```{r}

P8 <- ggplot(df_mainvariables, aes(sample=Depression))+
  stat_qq(color= "#D55E00", shapetype = "point", alpha = 0.5, bins = 15)+
  ggtitle("Depression QQ Plot") +
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5))+
  theme_classic()

P8
```


## 1.1.5.log10 transformation in Depression Score


```{r}

df_mainvariables<-  mutate (df_mainvariables, logDep = log10(Depression+1))

df_mainvariables


```


# 1.1.6.Check Normality (log10 Depression )


```{r}

P9 <- ggplot(data = df_mainvariables, aes(logDep)) +   # change position of title and axis label
  geom_histogram(aes(y = ..density..),bins = 15, position = "identity", color ="#D55E00", fill = "#D55E00",  linetype="dashed",  alpha = 0.5) + 
  ggtitle("Log10(Depression)Histogram and Density Graph") +
  geom_vline(aes(xintercept=mean(logDep)),
             color="#000000", linetype="dashed", size=0.75) +
  xlab("Log10(Depresyon)")+
  ylab ("Density") + 
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5)) +
  stat_function(fun = dnorm,
                args = list(mean = mean(df_mainvariables$logDep),
                            sd = sd(df_mainvariables$logDep)),
                col = "#000000",
                size = 1) + 
  theme_classic()

P9
```


```{r}

P10 <- ggplot(df_mainvariables, aes(sample=logDep))+
  stat_qq(color= "#D55E00", shapetype = "point", alpha = 0.5, bins = 15)+
  ggtitle("Log10(Depression) QQ Graph") +
  theme(legend.position =  "top", plot.title = element_text(hjust = 0.5))+
  theme_classic()

P10

```

## 2. Corelation Analysis 

```{r}

df_cor <-df_all %>%
        dplyr::select(BEEtop, RTOTOP, BIFOTOP, BDETOP)

df_cor <- mutate (df_cor, logDep = log10(BDETOP+1))


colnames(df_cor)[1] <- "CogFlex"
colnames(df_cor)[2] <- "Rumination"
colnames(df_cor)[3] <- "Mindfulness"
colnames(df_cor)[4] <- "Depression"
colnames(df_cor)[5] <- "LogDep"



MISSING <- 
           is.na(df_cor$CogFlex) |
           is.na(df_cor$Rumination) |
           is.na(df_cor$Mindfulness) |
           is.na(df_cor$Depression) |
           is.na(df_cor$LogDep) 


sum(MISSING) # Count the number of rows flagged for deletion

df_cordata<- subset(df_cor, 
                      subset = !MISSING) #Use ! to include those that are NOT missing


drop <- "Depression" # for mutlitple selection c("x","z")

df_cordata<-  df_cordata[,!(names(df_cordata) %in% drop)]


```


```{r}

cor (df_cordata, method = "pearson", use = "complete.obs") 

```


```{r}

GGally::ggpairs(df_cordata, columns = 1 : 4, 
                 lower = list(continuous = wrap("smooth", size =1, color ="blue")),
                 upper = list(continuous = wrap("cor", method = "pearson",size = 3, digits =2)),
                columnLabels = c("Cognitive Flexibility", "Rumination", "Mindfulness", "Depression"),
                 diag = list(continuous = "blankDiag")
                 )

```


```{r}
apa.cor.table(df_cordata, filename = "korelasiontable.doc", table.number = 1)
```

## 3. Mediatör Analyses 

```{r}
df_reg <-df_all %>% dplyr::select(CINSIYET,BEEtop, RTOTOP, BIFOTOP, BDETOP)

df_reg <- mutate (df_reg, logDep = log10(BDETOP+1))


colnames(df_reg)[1] <- "Gender" # female = 0 , male = 1 
colnames(df_reg)[2] <- "CogFlex"
colnames(df_reg)[3] <- "Rumination"
colnames(df_reg)[4] <- "Mindfulness"
colnames(df_reg)[5] <- "Depression"
colnames(df_reg)[6] <- "LogDep"




MISSING <- 
           is.na(df_reg$CogFlex) |
           is.na(df_reg$Rumination) |
           is.na(df_reg$Mindfulness) |
           is.na(df_reg$Depression) |
           is.na(df_reg$LogDep) 


sum(MISSING) # Count the number of rows flagged for deletion

df_regdata<- subset(df_reg, 
                      subset = !MISSING) #Use ! to include those that are NOT missing


drop <- "Depression" 

df_regdata<-  df_regdata[,!(names(df_regdata) %in% drop)]

```

#3.1. Assumption Check 

Y : Depression, X : Cognitive Flexibility, Mediatörs : Rumination (M1) and Mindfulness(M2), Covariance: Gender

```{r}

model_mediatör <- lm(LogDep ~  CogFlex + Rumination  + Mindfulness + Gender, data = df_reg )

summary(model_mediatör)

df_reg <- mutate (df_reg, logRum = log10(Rumination))


model_mediatör_t <- lm(LogDep ~  CogFlex + logRum  + Mindfulness + Gender, data = df_reg )
summary(model_mediatör)

# Model Diagnotistic 

# linearity 

car::crPlots(model_mediatör_t , terms = ~ CogFlex + logRum + Mindfulness,
             pch=20, col="gray",
             smooth = list(smoother=car::gamLine))

# residual normality

model_resid <- model_mediatör_t $residuals
length(model_resid) #423

length(coef(model_mediatör_t))-1 #4 

# There are greater than 10 observations per predictor (423 / 4 = 141).

par(mfrow=c(1,2))

hist(model_resid, xlab = "Residuals", probability = T,
     
     ylim = c(0, 0.6))

lines(density(model_resid, na.rm=T), lwd = 2, col = "red")
curve(dnorm(x, mean = mean(model_resid, na.rm=T), sd = sd(model_resid, na.rm=T)),
      lty = 2, lwd = 2, add = TRUE, col = "blue")

# Normal quantile-quantile (QQ) plot
# (use standardized residuals)
qqnorm(rstandard(model_mediatör_t), col="red", pch=20)
abline(a=0, b=1, col="blue", lty=2, lwd=2)

# Homoscedasticity or Constant Variance
car::residualPlots(model_mediatör_t,
                   pch=20, col="gray",
                   fitted = T,
                   ask = F, layout = c(3,2),
                   tests = F, quadratic = F)

#Check for Independence of Errors
car::durbinWatsonTest(model_mediatör_t) 

#Outliers and Influence Points 

car::outlierTest(model_mediatör_t )  # No Studentized residuals with Bonferroni p < 0.05

car::influenceIndexPlot(model_mediatör_t,
                        vars = c("Studentized", "hat"),
                        id=F,
                        main = "Residuals and leverage")

car::influenceIndexPlot(model_mediatör_t, vars = "Cook",
                        id=F, main = "Cook's distance")


#Multicollinearity 

car::vif(model_mediatör_t)  # VIF values close to 1 

```


# 3.2. Mediatör Analysis

```{r}

process(data = df_reg, y = "LogDep", x = "CogFlex", m= c("logRum", "Mindfulness"),
        model = 6, center = 2, moments = 1, modelbt = 1,total =1,cov = "Gender", stand = TRUE,  
        boot = 5000, seed =654321, conf = 95)

```





