---
title: "T-Test"
 author : "Kübra Mahmut Önalan"
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages  

```{r}

library(readxl) 
library (dplyr)
library(rstatix)
library(car) 
library(ggpubr)

```

# set directory

```{r}

setwd("")

```

# Import dataset and organize 

```{r}

df_all <- read_excel("")


df_mainvariables <- df_all %>%
  select(CINSIYET, RTOTOP)

df_mainvariables


colnames(df_mainvariables)[1] <- "Gender"
colnames(df_mainvariables)[2] <- "Rumination"

df_mainvariables

df_mainvariables$Gender[df_mainvariables$Gender == 1] <- 2 # male
df_mainvariables$Gender[df_mainvariables$Gender == 0] <- 1 # female

df_mainvariables$Gender <- factor(df_mainvariables$Gender ,levels = c(1,2), labels = c("Female", "Male"))

df_mainvariables


```

# Checking Missing Value 

```{r}


MISSING <- is.na(df_mainvariables$Rumination) 
           
           

sum(MISSING) 


df_mainvariables<- subset(df_mainvariables, 
                      subset = !MISSING) 



```



# Normality for Rumination by gender 

The normality assumption of rumination scores by gender was tested and it was decided that rumination scores were normally distributed by gender. 

```{r}

hist(df_mainvariables$Rumination[df_mainvariables$Gender=='Male'],main='Histogram for Male', xlab='Rumination')


```

```{r}

hist(df_mainvariables$Rumination[df_mainvariables$Gender=='Female'],main='Histogram for Female',xlab='Rumination') 


```


#Check mean and sd 

```{r}

mean_sd <- df_mainvariables %>% group_by(Gender) %>% get_summary_stats(Rumination, type = "mean_sd")

mean_sd
```

# Checking variance 

The Levene's test p-value=0.06 which is greater than 0.05 and so equality of variances for
the two independent samples can be assumed and the (student)t-test used.

```{r}

leveneTest(df_mainvariables$Rumination, df_mainvariables$Gender,center ="mean")

```


# 1.  T- test result 
```{r}

tresult <- t.test(Rumination ~ Gender,var.equal = TRUE, data = df_mainvariables)

tresult

```

# 2. Again T- test result, effect size and boxplot according to the result 

```{r}

stat.test <- df_mainvariables %>% 
  t_test(Rumination ~ Gender,var.equal = TRUE) %>%
  add_significance()

stat.test


df_mainvariables %>%  cohens_d(Rumination ~ Gender, var.equal = TRUE)


# Add p-value and significance levels
stat.test <- stat.test %>% add_xy_position(x = "Gender")

D<-ggboxplot(
  df_mainvariables, x = "Gender", y = "Rumination", 
  ylab = "Rumination", xlab = "Gender", add = "jitter",color = "Gender"
)+ 
  stat_pvalue_manual(stat.test, tip.length = 0) +
  labs(subtitle = get_test_label(stat.test, detailed = TRUE)) +
  scale_color_manual(values=c("#CC79A7", "#56B4E9", "#E69F00")) + 
  labs(color = "Gender") + 
  #theme_classic()
  theme_classic(base_size = 16, base_rect_size = 16, base_family = "Times New Roman")

D

ggsave("boxplotrumination.png")


```
